---
title: "GitHub"
author: "Kate Llangwig"
date: "4/20/2023"
output:
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Workshop Goals
- Introduce Git & GitHub
- learn to link GitHub to RStudio
- edit your first project and send it to GitHub

## What is Git?
- Git is version control system, with the original purpose of allowing groups to work collaboratively on software projects
- Git manages the evolution of a set of files - called a repository
- A repository is essentially a folder where you store your stuff
- Version control works a bit like "Track Changes" in word, Git will track the changes we make to our code so we can return to previous versions
- It also allows collaboration so I can look at your code and make changes - a bit like a more complicated version of Google Docs

## What is GitHub?
- GitHub is place where you 'share' your projects
- It is an online service (like google drive or Dropbox) for software
- Anyone can sign up for a free account
- GitHub is a for profit, but offers educational discounts


## Why use Git and GitHub?
- Git allows us to track changes to our code and projects in a systematic way
- We can restore our code at any point in time to fix errors we may have made
- It also allows for simpler sharing and merging of code across teams


## But I only code alone! 
- You need to carefully document your steps if the only person you are sharing code with is the future version of yourself

- In addition, most journals require publicly available data and code  - open code is the norm, not the exception.

- Using Git has gotten easier. We used to have to use command line to communicate with Git, but now we can just use RStudio! 

## Terminology

- repository: A directory or storage space where your projects can live. Sometimes GitHub users shorten this to “repo.” (If you're cool like that.) It is usually a local folder on your computer. You can keep code files, text files, image files, you name it, inside a repository.

- commit: This is the command that gives Git its power. When you commit, you are taking a “snapshot” of your repository at that point in time, giving you a checkpoint to which you can reevaluate or restore your project to any previous state.When you first start "commiting", it is important to remember this is taking the picture, not SENDING the picture. (Sending is called "pushing")

## Terminology cont.

- branch: How do multiple people work on a project at the same time without Git getting them confused? Usually, they “branch off” of the main project with their own versions full of changes they themselves have made. After they’re done, it’s time to “merge” that branch back with the “master,” the main directory of the project. Because we'll be working within our own repos, we don't need to worry too much about branching but is good to know for future.

- push: This is how you upload your file to GitHub. Remember, you need to both commit and push for your file to be sent to GitHub. 

## Command line vs Desktop Client
- there are numerous programs (like GitHub Desktop) that can make working with Git and GitHub easier
- We can work with Git using the command line (e.g. in the 'terminal') which is how it used to HAVE to be done
- IMO, using tools like RStudio and the Desktop Client make Git more accessible, particularly for beginners
- there is no nerd olympics! It's okay not to use the command line for everything. 

# Pre-course homework

## Register for a GitHub username
- https://github.com/
- click on either of the big green sign up buttons

## Installing Git
 - Hopefully you have installed Git!
 - If not, please do so here, following the instructions:
 http://happygitwithr.com/install-git.html#install-git
 - Try to install git in the most scientific way possible - if one way doesn't work, try the next, and google your mistakes! 
 
## Install a Git client
- GitHub desktop is great for beginners
- https://desktop.github.com

## Enter your user name into this form
- enter your github user name [here](https://docs.google.com/forms/d/e/1FAIpQLSfH3dHSr-UuiPo84EH_4smshl9fvJ-t7x2hA30Dx8dcm0Hy0w/viewform?usp=sf_link)
 
 
## Tasks for today
- Configure Git on your computer
- Create a personal access token
- Set up a repository on GitHub
- clone that repository to your computer
- edit something in your repository
- commit your edits using Git
- Push your edits to GitHub

## 1. Configure Git on your computer
- Open RStudio & click File-->New File --> RScript
- Type the following:
```{r, eval=FALSE, echo=TRUE}
## install if needed (only do this once)
## install.packages("usethis")
library(usethis)
use_git_config(user.name = "Jane Doe", "user.email = jane@vt.edu")
```
- Make sure you use the user name and email you used to sign up for github!  

## 2. Create a personal access token
- when we interact with Git, we need to tell it that we are a specific user that is allowed to make changes to the folder
- We have to create a token that we store on our computers to allow us access to connect to GitHub
- open 
